%
% snowstar.tex -- 
%
% (c) 2016 Prof Dr Andreas Mueller, Hochschule Rapperswil 2016
%
\chapter{Snowstar: the network daemon\label{chapter:snowstar}}
The \texttt{snowstar} daemon is designed to be run on a small
computer mounted directly on the telescope or the mount, and to
control cameras and other devices.
Clients then don't need any knowledge of the intricacies of controlling
hardware devices.

\section{Running the daemon}
The daemon understands the following options:
\begin{itemize}
\item
\texttt{-b,--base=}\textit{imagedir}
\\
The daemon needs a directory where it can temporarily store images until
a client retrieves them and removes them from the daemon's scope.
\item
\texttt{-c,--config}\textit{configdb}
\\
The daemon uses a database for configuration information.
\item
\texttt{-d,--debug}
Increase the debug level.
The debug messages are written to standard output.
\item
\texttt{-D,--database=}\textit{databasefile}
\\
The daemon uses a database to persist data for temporary objects
like tasks in the task queue, instrument configuration, guider
calibrations and guide protocols.
In most cases, the same database is used for configuration and
for persistence.
\item
\texttt{-h,--help}
\\
Display a help message and exist.
\item
\texttt{-f,--foreground}
\\
In normal operation, the daemon goes into the background.
For debugging, it is sometimes more convenient to have it run in the
foreground, which can be forced with this option.
\item
\texttt{-g,--group=}\textit{group}
\\
Change group id to the one of~\textit{group} at run time.
This is most useful when starting the daemon from a start script
that runs as \texttt{root}.
This option must be given before the database options to ensure
that the process runs under the right group id when accessing the
databases.
\item
\texttt{-l,--logfile=}\textit{logfile}
\\
Write log messages to a log file named \textit{logfile} instead of
standard error.
\item
\texttt{-p,--port=}\textit{port}
\\
Run the service on this port.
There also is a configuration option to configure the TCP port to use
for the service, but this option might be handy during debugging.
\item
\texttt{-P,--pidfile=}\textit{pidfile}
\\
Full path name to the pid file to create when starting up.
The pid file can be used by stop-scripts to get the process id
to of the currently running process.
\item
\texttt{-n,--name=}\textit{servicename}
\\
Run the service with the Zeroconf name \textit{servicename}.
The network clients will use this service name to locate the service.
\item
\texttt{-s,--sslport=}\texttt{sslport}
\\
Run the SSL service on a separate port.
As for the unencrypted port, there is a configuration option to set the
SSL port which is usually preferred.
\item
\texttt{-L,--syslog}
\\
Send log to syslog instead of standard error.
\item
\texttt{-u,--user=}\textit{user}
\\
Change user id to the one of~\textit{user} at run time.
This is most useful when starting the daemon from a start script
that runs as \texttt{root}.
This option must be given before the database options to ensure
that the process runs under the right group id when accessing the
databases.

\end{itemize}

\section{Services offered by the \texttt{snowstar} server}
The \texttt{snowstar} server offers all services needed for an
astrophotography workflow.
In most cases, not all services will be needed on all systems.
One will usually restrict the database of instruments to a single
host, and image repositories to a host that actually has enough
storage to provide the image repository service.

The services are identified by a TXT record in ZeroConf multicast DNS.
Here is what the services offer:
\begin{itemize}
\item \texttt{devices}: Access to hardware components attached to the
machine offering the service.
\item \texttt{images}: CCD devices allow taking images, and which a 
client may want to inspect.
The client has to use the \texttt{images} service for this purpose.
If a service is activated that may require the use of the \texttt{images}
service, the service will be available, but it will not be published in
ZeroConf.
\item \texttt{instruments}: The \texttt{instruments} service is merely
a database service to inform other services of the available instruments,
their components and some parameters of interest.
\item \texttt{tasks}: The \texttt{tasks} service runs preprogrammed
sequences of observations. It should normally be run on the machine that
controls the camera used for the images.
\item \texttt{guiding}: The \texttt{guiding} service performs guiding,
it is preferably run on the machine managing the guide camera, but
it can in principle be run on any machine.
\item \texttt{focusing}: This service allows to automatically focus a
camera/focuser combination.
\item \texttt{repository}: This service makes the content of image
repositories available.
Note that it is possible to use image repositories e.~g.~to store images
taken by the \texttt{guiding} and \texttt{focusing} services, even if
the this service is not enabled.
The server will store the images in the repositories, but it will not
be possible to access the image via the network.
\end{itemize}


\section{Configuration}
The server can be configured to publish only a subset of the available
services, by setting the following global configuration variables according
to table~\ref{snowstar:configuration:table}.
\begin{table}
\centering
\begin{tabular}{lll}
section         &servicename         &default     \\
\hline
\texttt{service}&\texttt{devices}    &\texttt{yes}\\
\texttt{service}&\texttt{images}     &\texttt{yes}\\
\texttt{service}&\texttt{instruments}&\texttt{no} \\
\texttt{service}&\texttt{tasks}      &\texttt{no} \\
\texttt{service}&\texttt{guiding}    &\texttt{no} \\
\texttt{service}&\texttt{focusing}   &\texttt{no} \\
\texttt{service}&\texttt{repository} &\texttt{no} \\
\hline
\end{tabular}
\caption{The configuration variables used to configure the services
offered by the snowstar server.
\label{snowstar:configuration:table}}
\end{table}




